Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-01-20 2_SSL_EES/article/article_main.tex
updated figure caption

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines. Dashed lines indicate the average improvement calculated from a 3rd-order polynomial fit to the data. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID – High-Intensity Discharge; CFL – Compact Fluorescent Lamp; Hal. – Halogen, Incd. – Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, section , for the full list of data sources. }
 \label{fgr:history_efficacy}
\end{figure*}

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies}\cite{Anadon2016}. Among these drivers, the role of external  knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area  \cite{Stephan2021}. While the exact definition of spillovers in literature depends on the context and the research question \cite{Liu2003}\cite{Nemet2012}, we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external – thus a spillover –  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be an important factor in understanding \cite{Pichler2020} and shaping \cite{Clark2016}\cite{Stephan2021}\cite{Sun2021}\cite{kolesnikov_framework_2022} the future evolution of technologies.

@@

\begin{figure}
\centering
  \includegraphics[height=6.5cm]{2_SSL_EES/article/figures/cost_lamp_small.pdf}
  \caption{Historical development of retail sales prices (in 2020 USD) per luminous flux of LED-based luminaires, including light bulbs, spotlights and recessed lights from 2008 to 2020. Red curved and dashed lines represent average retail sales prices and price projections for LED based luminaires published by the U.S. Department of Energy (DoE) \cite{council2013assessment}. Shown for reference are the average prices for compact fluorescent (CFL) and incandescent light bulbs. Source: own synthesis of data on LED sales prices collected from various consumer watchdog databases and publications. Data on CFL sales prices collected from \cite{eger2018origin} as well as various consumer watchdog databases and publications. Incandescent light bulb price is assumed constant based on the average in the covered time period. See Supplementary information, section …, for the full list of data sources.}
  \label{fgr:cost_lamp_small}
\end{figure}

@@

\begin{figure*}
 \centering
 \includegraphics[height=17cm]{2_SSL_EES/article/figures/chip_architecture_overview.pdf}
 \caption{Historical evolution of light-emitting diode chip architectures. Rows from top to bottom: classical chip with lateral current spreading; Osram’s thin GaN flip-chip (vertical) architecture; flip-chip architecture; chip-scale package flip-chip architecture. Shown are side views of chips without packages, along a cutaway line best suited to the features of each architecture. The cutaway surface is indicated by a red arrow with an eye on the overlaid top view of each chip. Note that the dimensions are not to scale, and smaller features are greatly exaggerated for clarity. Years indicated correspond to the earliest identified patent priority date. Black frames around certain designs indicate chip designs not brought to large-scale production. Abbreviations: GaN – Gallium Nitride, ITO – Indium Tin Oxide, MQW – Multiple Quantum Well. Source: adapted and compiled from multiple patents and industry publications. See Supplementary Information, section…, for the full list of sources and references.}
 \label{fgr:chip_architecture_overview}
\end{figure*}

@@

Device forward voltage efficiency\footnote{"Joule Efficiency" $\epsilon_J$ in Tsao et al. \cite{tsao2010solid}} $V_fE$ describes all electrical losses at the interface between the electrodes and the semiconductor and in the bulk. These losses can be due to tunneling and Ohmic resistance at the interface, as well as Ohmic resistance and other electrical losses within the bulk of the semiconductor. It is defined as

@@

Device light-extraction efficiency describes the  losses due to absorption in the material after electron-hole recombination and the associated emission of a photon. It is defined as

@@

Internal quantum efficiency describes non-radiative recombinatory processes in the semiconductor bulk. It is related to the external quantum efficiency through
\begin{equation}
\label{eqn:ieq-eqe}
    \eta_{EQ} = \eta_{IQ} \times \eta_{LE}
\end{equation}

@@

(Light) conversion efficiency describes losses in the conversion process of blue light that is the basis for all phosphor-converted white LEDs. These losses are the sum of the Stokes loss and well as scattering/absorption losses. It is defined as

@@

Spectral efficiency describes losses in the conversion process due to the wavelength-dependent efficiency of the human eyes. Photons converted into the infrared or ultraviolet are lost to illumination purposes. It is defined as

@@

The Colour Rendering Index (CRI) of a light source describes its ability to render the colours of an object faithfully when compared to illumination under a reference light source, such as standard daylight \cite{khan2015led}. The way it is calculated is defined by the International Illumination Commission (CIE) \cite{cie_cri_1995}. CRI has certain limitations when applied to solid-state light sources \cite{david2013cri}. However, despite repeated attempts at constructing more elaborate colour rendering metrics \cite{Houser2013}, CRI has remained the de-facto industry standard for describing colour fidelity of light sources \cite{DoE2016LED}. High colour rendering performance of lighting is a requirement in workplace environments, retail stores, clinical operating environments and art exhibitions \cite{khanh2017color}. It should be noted that some niche applications prioritize high colour saturation over high CRI, for instance, in food display or fabric retail applications \cite{david2013cri}. However, these niche applications remain outside our focus on general illumination. Due to a broad availability and importance of CRI data for consumers of various LED lighting sources, we adopt CRI as the key metric to track progress in consumer experience in LED lighting, despite its limitations.

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures from Figure \ref{fgr:chip_architecture_overview} are shown for reference: 2003 (top panel) – flip-chip with YGAG phosphor; 2010 (bottom panel) – flip-chip with 258 phosphor; 2016 (top panel) – flip-chip with PFS phosphor; 2020 (bottom panel) – flip-chip with SALON phosphor. Details on phosphors are provided in Table \ref{tab:phosphors} and SI section XX. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003 and $\eta_L = 12.7\%$ in 2010. Abbreviations: Scat. = Scattering. Source: own elaboration based on data from sources listed in SI, section XX.}
 \label{fgr:waterfall}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_efficiency.pdf}
 \caption{Contribution of innovations and technology spillovers to the historical progress in sub-efficiencies of phosphor-converted warm white LEDs with test currents of at least 350mA. Vertical bars represent LED technology innovations, with purple bars indicating innovations driven by technology spillovers (annotated and listed in an inset table) and grey bars indicating all other improvements identified in this study. Horizontal coloured lines indicate state-of-the-art sub-efficiency levels for five years: 1997, 2003, 2010, 2016, and 2020. Physical limits on sub-efficiencies are indicated by black horizontal lines. The overall LED lamp efficiency, $\eta_L$, is displayed in the rightmost column. $\eta_Vf$ – forward voltage efficiency; $\eta_IQ$ – internal quantum efficiency; $\eta_Droop$ – efficiency droop; $\eta_LE$ – light extraction efficiency; $\eta_(C_ER)$ – conversion efficiency for red phosphors; $\eta_(C_(EY/G))$ – conversion efficiency for yellow/green phosphors; $\eta_S$ – spectral efficiency; PSS – patterned sapphire substrate, ITO – indium tin oxide current spreading layer; 258, SLA, PFS – different phosphors used in white LEDs, see Table 3. "N/A" denotes sub-efficiencies that could not be calculated for 1997 due to $\eta_Vf$, $\eta_droop$ depending on current, which was below 350mA at the time, and warm white spectrum LEDs not available at the time for $\eta_(C_ER)$,$\eta_(C_(EY/G))$  and $\eta_S$ The physical limit on $\eta_Vf$ exceeds 100\% and depends on electrical device parameters \cite{david2016electrical}. Source: own elaboration based on data from sources listed in SI, section XX.}
 \label{fgr:breakthroughs_efficiency}
\end{figure*}

@@

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Circle dots show data for the earliest identified representative white LED products with published spectral data that used phosphors listed in Table 3, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI1 in SI section 2. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. Desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG – Nichia, 1998 \cite{bando1998development}; YGAG – Lumileds, 2002 \cite{Mueller2002}; 258 – Lumileds, 2005 \cite{MuellerMach2005}; SLA – Lumileds, 2014 \cite{Pust2014}; PFS – GE, 2015 \cite{Murphy2015}; QD – Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON – Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

>>>>

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/history_efficacy.pdf}
 \caption{Historical development of the luminous efficacy ($K$) of the most widely-used lighting technologies in human history from 1850 to 2020. Data shown for best performers by year of market introduction, with dashed lines. Dashed lines indicate the average improvement calculated from a 3rd-order polynomial fit to the data. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_max^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID - High-Intensity Discharge; CFL - Compact Fluorescent Lamp; Hal. - Halogen, Incd. - Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Information, section , for the full list of data sources. }
 \label{fgr:history_efficacy}
\end{figure*}

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies}\cite{Anadon2016}. Among these drivers, the role of external  knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area  \cite{Stephan2021}. While the exact definition of spillovers in literature depends on the context and the research question \cite{Liu2003}\cite{Nemet2012}, we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external - thus a spillover -  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be an important factor in understanding \cite{Pichler2020} and shaping \cite{Clark2016}\cite{Stephan2021}\cite{Sun2021}\cite{kolesnikov_framework_2022} the future evolution of technologies.

@@

\begin{figure}
\centering
  \includegraphics[height=6.5cm]{2_SSL_EES/article/figures/cost_lamp_small.pdf}
  \caption{Historical development of retail sales prices (in 2020 USD) per luminous flux of LED-based luminaires, including light bulbs, spotlights and recessed lights from 2008 to 2020. Red curved and dashed lines represent average retail sales prices and price projections for LED based luminaires published by the U.S. Department of Energy (DOE) \cite{council2013assessment}. Shown for reference are the average prices for compact fluorescent (CFL) and incandescent light bulbs. Source: own synthesis of data on LED sales prices collected from various consumer watchdog databases and publications. Data on CFL sales prices collected from \cite{eger2018origin} as well as various consumer watchdog databases and publications. Incandescent light bulb price is assumed constant based on the average in the covered time period. See Supplementary information, section …, for the full list of data sources.}
  \label{fgr:cost_lamp_small}
\end{figure}

@@

\begin{figure*}
 \centering
 \includegraphics[height=17cm]{2_SSL_EES/article/figures/chip_architecture_overview.pdf}
 \caption{Historical evolution of light-emitting diode chip architectures. Rows from top to bottom: classical chip with lateral current spreading; Osram’s thin GaN flip-chip (vertical) architecture; flip-chip architecture; chip-scale package flip-chip architecture. Shown are side views of chips without packages, along a cutaway line best suited to the features of each architecture. The cutaway surface is indicated by a red arrow with an eye on the overlaid top view of each chip. Note that the dimensions are not to scale, and smaller features are greatly exaggerated for clarity. Years indicated correspond to the earliest identified patent priority date. Black frames around certain designs indicate chip designs not brought to large-scale production. Chip architecture abbreviations: TF - Thin-Film; FC - Flip-Chip; CSP - Chip-Scale Package; Material abbreviations: GaN - Gallium Nitride, ITO - Indium Tin Oxide, MQW - Multiple Quantum Well. Source: adapted and compiled from multiple patents and industry publications. See Supplementary Information, section…, for the full list of sources and references.}
 \label{fgr:chip_architecture_overview}
\end{figure*}

@@

Device forward voltage efficiency\footnote{"Joule Efficiency" $\epsilon_J$ in Tsao et al. \cite{tsao2010solid}} (VfE) describes all electrical losses at the interface between the electrodes and the semiconductor and in the bulk. These losses can be due to tunneling and Ohmic resistance at the interface, as well as Ohmic resistance and other electrical losses within the bulk of the semiconductor. It is defined as

@@

Device light-extraction efficiency (LEE) describes the  losses due to absorption in the material after electron-hole recombination and the associated emission of a photon. It is defined as

@@

Internal quantum efficiency (IQE) describes non-radiative recombinatory processes in the semiconductor bulk. It is related to the external quantum efficiency through
\begin{equation}
\label{eqn:ieq-eqe}
    \eta_{EQ} = \eta_{IQ} \times \eta_{LE}
\end{equation}

@@

(Light) conversion efficiency (LCE) describes losses in the conversion process of blue light that is the basis for all phosphor-converted white LEDs. These losses are the sum of the Stokes loss and well as scattering/absorption losses. It is defined as

@@

Spectral efficiency (SE) describes losses in the conversion process due to the wavelength-dependent efficiency of the human eyes. Photons converted into the infrared or ultraviolet are lost to illumination purposes. It is defined as

@@

The Colour Rendering Index (CRI) of a light source describes its ability to render the colours of an object faithfully when compared to illumination under a reference light source, such as standard daylight \cite{khan2015led}. The way it is calculated is defined by the International Illumination Commission (CIE) \cite{cie_cri_1995}. CRI has certain limitations when applied to solid-state light sources \cite{david2013cri}. However, despite repeated attempts at constructing more elaborate colour rendering metrics \cite{Houser2013}, CRI has remained the de-facto industry standard for describing colour fidelity of light sources \cite{DOE2016LED}. High colour rendering performance of lighting is a requirement in workplace environments, retail stores, clinical operating environments and art exhibitions \cite{khanh2017color}. It should be noted that some niche applications prioritize high colour saturation over high CRI, for instance, in food display or fabric retail applications \cite{david2013cri}. However, these niche applications remain outside our focus on general illumination. Due to a broad availability and importance of CRI data for consumers of various LED lighting sources, we adopt CRI as the key metric to track progress in consumer experience in LED lighting, despite its limitations.

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2003.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2010.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2016.pdf}
 \includegraphics[width=17cm]{2_SSL_EES/article/figures/waterfall_performance_2020.pdf}
 \caption{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020 (top to bottom panels), normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Section \ref{sec:device_performance_metrics}. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for red, green and blue loss channels indicate losses of remaining red/green/blue light energy. The corresponding LED architectures from Figure \ref{fgr:chip_architecture_overview} are shown for reference: 2003 (top panel) - flip-chip with YGAG phosphor; 2010 (bottom panel) - flip-chip with 258 phosphor; 2016 (top panel) - flip-chip with PFS phosphor; 2020 (bottom panel) - flip-chip with SALON phosphor. Details on phosphors are provided in Table \ref{tab:phosphors} and SI section XX. Overall LED package efficiency is $\eta_L = 5.8\%$ in 2003 and $\eta_L = 12.7\%$ in 2010. Abbreviations: Scat. = Scattering. Source: own elaboration based on data from sources listed in SI, section XX.}
 \label{fgr:waterfall}
\end{figure*}

@@

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_efficiency.pdf}
 \caption{Contribution of innovations and technology spillovers to the historical progress in sub-efficiencies of phosphor-converted warm white LEDs with test currents of at least 350mA. Vertical bars represent LED technology innovations, with purple bars indicating innovations driven by technology spillovers (annotated and listed in an inset table) and grey bars indicating all other improvements identified in this study. Horizontal coloured lines indicate state-of-the-art sub-efficiency levels for five years: 1997, 2003, 2010, 2016, and 2020. Physical limits on sub-efficiencies are indicated by black horizontal lines. The overall LED lamp efficiency, $\eta_L$, is displayed in the rightmost column. $\eta_Vf$ - forward voltage efficiency; $\eta_IQ$ - internal quantum efficiency; $\eta_Droop$ - efficiency droop; $\eta_LE$ - light extraction efficiency; $\eta_(C_ER)$ - conversion efficiency for red phosphors; $\eta_(C_(EY/G))$ - conversion efficiency for yellow/green phosphors; $\eta_S$ - spectral efficiency; PSS - patterned sapphire substrate, ITO - indium tin oxide current spreading layer; 258, SLA, PFS - different phosphors used in white LEDs, see Table 3. "N/A" denotes sub-efficiencies that could not be calculated for 1997 due to $\eta_Vf$, $\eta_droop$ depending on current, which was below 350mA at the time, and warm white spectrum LEDs not available at the time for $\eta_(C_ER)$,$\eta_(C_(EY/G))$  and $\eta_S$ The physical limit on $\eta_Vf$ exceeds 100\% and depends on electrical device parameters \cite{david2016electrical}. Source: own elaboration based on data from sources listed in SI, section XX.}
 \label{fgr:breakthroughs_efficiency}
\end{figure*}

\subsection{Improvements in Sub-Efficiencies and Overall Efficiency}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=17.5cm]{2_SSL_EES/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. \newline
 \textbf{Panel A1:} Device forward voltage at a test current of I=350mA. The physical limit for a blue light wavelength of 450nm without electric pumping is shown for reference. Data points for devices released in 2020 by various manufacturers are shown in an inset plot.
 \textbf{Panel A2:} Efficiency droop at the test current of I=350mA.
 \textbf{Panel B1:} Internal quantum efficiency, for different chip architectures, by type of measurement used. Measurement methods: PL - Photoluminescence\cite{Shim_2018}, EL - Electroluminescence\cite{Getty_2009};  ABC Model\cite{Karpov_2014}, Ideality Factor\cite{Masui_2010}. “Not reported” denotes data points where the IQE measurement technique was not reported. Whiskers indicate reported ranges, where applicable.
 \textbf{Panel B2:} Light extraction efficiency for different chip architectures, by type of specific technology used to improve light-extraction efficiency. Abbreviations: TFFC - Thin-Film Flip-Chip; FC - Flip-Chip; CSP - Chip-Scale Package; ITO - Indium Tin Oxide; ZnO - Zinc Oxide; PSS - Patterned Sapphire Substrate; electr. - electrode; text. - textured.
 Source: own synthesis of data from the full list of sources provided in the Supplementary Information, section XX. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports\cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report\cite{beale_leds_2015}.}
 \label{fgr:waterfall}
\end{figure*}

@@

\begin{figure*}[h]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Circle dots show data for the earliest identified representative white LED products with published spectral data that used phosphors listed in Table 3, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI1 in SI section 2. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. Desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}
